### 2023-01-23 chaps/pol.tex
Use  citation style as default

This chapter will describe the polarimetric observation, experiments, and applications. A test long text. A test long text. A test long text. A test long text. A test long text. A test long text. A test long text. A test long text. A test long text. A test long text. A test long text. A test long text. A test long text. A test long text. A test long text. A test long text.

\cite{1929PhDT.........9L,1955PhDT........37D,1970GeCAS...1.2127G,1986MNRAS.218...75G}

Inside the paretheses \citep{2016MNRAS.456..248C}