### 2023-01-26 chaps/pol.tex
slight modification:

This chapter will describe the polarimetric observation, experiments, and applications. A test long text. A test long text. A test long text. A test long text. A test long text. A test long text. A test long text. A test long text. A test long text. A test long text. A test long text. A test long text. A test long text. A test long text. A test long text. A test long text.