### 2023-08-15 chaps/ch-emp.tex
publish rc

The reference in a PhD thesis can easily go beyond 100 items. The management will be tricky if you do not plan beforehand.

\section{Reference}
\subsection{ADS}
For astronomers, NASA provides a handy tool called ADS\footnote{\url{https://ui.adsabs.harvard.edu/}}, and its functionality, called ``Library'', is fantastic (though still in need of improvement). To use it efficiently, I have written a simple script called ``\texttt{ads2bibtex}''\footnote{\url{https://github.com/ysBach/ads2bibtex}}.

\subsection{Reference Style}\label{ss:ref style}
The reference style is nothing more than personal taste. Here I used the ``URL-ed A\&A style''\footnote{\url{https://github.com/yangcht/AA-bibstyle-with-hyperlink}}'' specified in the ``\texttt{aa\_url.bst}'' file. The \textcolor{magenta}{magenta link} redirects to the URL (usually DOI), and \textcolor{blue}{blue link} redirects to the ADS URL.
\textcolor{red}{I tweaked it a bit so that the bibliography shows the title of the publication in double quotes}.

\textbf{IMPORTANT NOTE}: As of 2023-05-18, at least \texttt{@PHDTHESIS} must have \texttt{adsurl=\{\}} and \texttt{adsnote=\{\}} if they are not registered with ADS. If you don't have these fields, you will get the following error:

  \texttt{Extra \}, or forgotten $\backslash$endgroup.}

\subsection{\texttt{cite}}\label{ss:cite}
You may already be familiar with citation. For your information:
\begin{itemize}[itemsep=-5pt, topsep=0pt, partopsep=0pt]
\item \texttt{cite} makes \cite{2021A&A...654A.113B}. If two arguments, \cite{2021A&A...654A.113B,2022MNRAS.509.4128I}.
\item \texttt{citet} makes \citet{2021A&A...654A.113B}. If two arguments, \citet{2021A&A...654A.113B,2022MNRAS.509.4128I}.
\item \texttt{citep} makes \citep{2021A&A...654A.113B}. If two arguments, \citep{2021A&A...654A.113B,2022MNRAS.509.4128I}.
\item \texttt{citealt} makes \citealt{2021A&A...654A.113B}. If two arguments, \citealt{2021A&A...654A.113B,2022MNRAS.509.4128I}.
\end{itemize}
With optional arguments, ``\texttt{$\backslash$citep[See, for example,][Figure 1]\{2021A\&A...654A.113B\}}'' becomes: \citep[See, e.g.,][Figure 1]{2021A&A...654A.113B}.

\section{\texttt{Cref}}\label{ss:cref}
\texttt{Cleveref} package provides multiple options for simple referencing. Please refer to the original documentation\footnote{\url{https://ctan.org/pkg/cleveref?lang=en}}.

\section{Code}
In this thesis, you can use \texttt{lstlisting} environment:

\begin{lstlisting}
# import nicpolpy package
import nicpolpy as nic

# Do not print useless warnings
import warnings
from astropy.utils.exceptions import AstropyWarning
warnings.filterwarnings('ignore',
    append=True, category=AstropyWarning)

# Cell 0: Initialize the reducer
npr = nic.NICPolReduc(
    name="SP_20190417",
    inputs="_original_32bit/190417/raw/*.fits",
    mflats="cal-flat_20180507-lv1/*.fits",
    imasks="masks/*.fits",
    verbose=1
)
\end{lstlisting}

You may modify the preference settings at the preamble in the ``\texttt{manuscript.tex}'' file.