### 2023-08-15 chaps/paper1.tex

\section{Summary}
Summary of this paper. In the natural sciences, it seems to be customary to include published papers in the dissertation. Especially if the paper has already been published, you can (i) add a footnote to the chapter title to indicate the original publication and (ii) put the abstract of the paper in this ``Summary'' or ``Abstract'' section. You can just paste the content into the next sections.

\section{Future Work}
You can put the content of the published paper here. If you want, you can add additional comments, explanations, figures, tables, etc. that were not included in the original publication.