### 2023-08-15 title.tex

% Thesis title
\def\tTitleA{{On the PhD Dissertation}}
\def\tTitleB{{at Seoul National University}}
\def\tTitleK{{서울대학교의 이학박사학위 \\학위논문에 관하여}}

% Department name
\def\tDepartment{{Department of Physics and Astronomy}}  % Department name
\def\tDepartmentKr{{물리$\cdot$천문학부 천문학전공}}           % Department name
\def\tCollegeOrGradSchool{{College of Natural Sciences}}  % Exists only in English format.. I dunno why.
\def\tMajor{{Astronomy}}  % Normally it is in separate line in English version. If you have major, add this. Otherwise, modify below (cosmetically modify as you wish)

% Submission date.
% I have no idea about official regulation, but mostly people do 6월 or 12월.
\def\tDate{{June 2023}}
\def\tDateKr{{2023년 6월}}

% Only February or August
\def\tInjunDate{{August 2023}}
\def\tInjunDateKr{{2023년 8월}}  % xxxx년 2월 OR 8월

% Name of the degree
\def\tDeg{{Ph.D.}}
\def\tDegKr{{이학박사}}  % 이학석사, 이학박사, 공학석사, 공학박사, ...
\def\tDegLong{{Doctor of Philosophy}}

% My name
\def\tName{{Gildong X. Hong}}
\def\tNameWithBlank{{Gildong\ X.\ Hong}}
\def\tNameKr{{홍길동}}  % If you want, you can put English name here too (I did it with my PhD thesis)
\def\tNameKrWithBlank{{홍\ 길\ 동}}  % If you want, you can put English name here too (I did it with my PhD thesis)

% Supervisor name
\def\tNameSuper{{Baksa\ F.\ Kim}}
\def\tNameSuperKr{{김\ 박\ 사}}   % If you want, you can put English name here too (I did it with my PhD thesis)

\newcommand{\wideunderline}[2][2em]{%
  \underline{\makebox[\ifdim\width>#1\width\else#1\fi]{#2}}%
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Korean Version
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\renewcommand{\baselinestretch}{1}
\centering
{\Large \quad \par}
\vspace{.5cm}{\Large \bf \tDegKr\ 학위논문 \par}
\vspace{2cm}
{\Huge \bf \tTitleA \par}
\vspace{.3cm}
{\Huge \bf \tTitleB \par}
\vspace{2cm}
{\LARGE \bf \tTitleK \par}
\vspace{4.5cm}
{\Large \bf \tInjunDateKr \par}
\vspace{3cm}
{\LARGE \bf 서울대학교 대학원 \par}  % I think it should be "서울대학교 대학원" all the time.
\vspace{.5cm}
{\Large \bf \tDepartmentKr \par}
\vspace{.5cm}
{\LARGE \bf \tNameWithBlank \par}

\clearpage

\centering
{\Large \quad \par}
\vspace{.1cm}
{\Huge \bf \tTitleA \par}
\vspace{.3cm}
{\Huge \bf \tTitleB \par}
\vspace{0.5cm}
{\LARGE \bf  \tTitleK \par}
\vspace{1.cm}
{\Large \bf 지도교수 $\ $ \tNameSuperKr  \par}
\vspace{0.5cm}
{\LARGE \bf 이 논문을 \tDegKr\ 학위논문으로 제출함 \par}
\vspace{.2cm}
{\Large \bf \tDateKr \par}

\vspace{.9cm}
{\LARGE \bf 서울대학교 대학원 \par}
\vspace{.3cm}
{\Large \bf \tDepartmentKr \par}
\vspace{.3cm}
{\LARGE \bf \tNameKrWithBlank \par}
\vspace{.9cm}

{\LARGE \bf \tNameKr 의 \tDegKr\ 학위논문을 인준함 \par}
\vspace{.2cm}
{\Large \bf \tInjunDateKr \par}
\vspace{1.2cm}
{\Large \bf  위\hspace{0.48cm}원\hspace{0.48cm}장} \quad {\Large \wideunderline[12em]{\hfill Professor N. One\hfill(인)}} \par
\vspace{.7cm}
{\Large \bf 부\ 위\ 원\ 장} \quad {\Large \wideunderline[12em]{\hfill Likely Ur. Supervisor\hfill(인)} } \par
\vspace{.7cm}
{\Large \bf 위\hspace{1.55cm}원} \quad {\Large \wideunderline[12em]{\hfill Another Professor\hfill(인)} } \par
\vspace{.7cm}
{\Large \bf  위\hspace{1.55cm}원} \quad {\Large \wideunderline[12em]{\hfill Another Doctor\hfill(인)} } \par
\vspace{.7cm}
{\Large \bf 위\hspace{1.55cm}원} \quad {\Large \wideunderline[12em]{\hfill Yet Another\hfill(인)} } \par

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% An Information page in English
%% I made it by myself following few predecessors. I don't know if this is an official format.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage
\centering

{\Large \quad \par}
\vspace{.1cm}
{\Huge \bf \tTitleA \par}
\vspace{.3cm}
{\Huge \bf \tTitleB \par}
\vspace{.5cm}
{\large By \par}
\vspace{0.5cm}
{\Large \bf \tNameWithBlank \par }
\vspace{0.5cm}
{\large A dissertation submitted in partial fulfillment of the requirements \\ for the degree of \par}
\vspace{.5cm}
{\Large \bf \tDegLong \par}
\vspace{.3cm}
{\large in\par}
\vspace{.3cm}
{\Large \tMajor \par}
\vspace{.9cm}

{\large \tCollegeOrGradSchool \par}
\vspace{.2cm}
{\large \tDepartment \par}
\vspace{.2cm}
{\large Seoul National University\par}
\vspace{.2cm}
{\large \tMajor\ Major \par}

\vspace{1cm}
{\large Supervised by $\ $ \tNameSuper \par}
\vspace{1cm}

{\large \tInjunDate \par}
\vspace{1.2cm}
{\large Committee: \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad \par}

\begin{table} [h!]
\large
\centering
\begin{tabular}{lll}\vspace{.2cm}
Chair      & Professor & Professor N. One \\ \vspace{.2cm}
Vice Chair & Professor & Likely Ur. Supervisor \\ \vspace{.2cm}
Examiner   & Professor & Another Professor\\ \vspace{.2cm}
Examiner   & Doctor & Another Doctor\\ \vspace{.2cm}
Examiner   & Doctor & Yet Another\\
\end{tabular}
\end{table}
\clearpage